### 2019-12-18 paper/contrib.tex

% !TEX root = nanograv_11yr_cw.tex
This document is the result of more than a decade of work by the entire NANOGrav collaboration.
We acknowledge specific contributions below.
Z.A., K.C., P.B.D., M.E.D., T.D., J.A.E., R.D.F., E.C.F., E.F., P.A.G., G.J., M.L.J., M.T.L., L.L., D.R.L., R.S.L., M.A.M., C.N., D.J.N., T.T.P., S.M.R., P.S.R., R.S., I.H.S., K.S., J.K.S., and W.W.Z. developed the 11-year data set.
M.V.\ led this analysis and coordinated the paper writing.
M.V., S.R.T., and J.S. developed \textsc{BayesEphem}, managed Bayesian-inference runs, and performed data analysis, with help from C.C., J.A.E., T.J.W.L., and S.J.V. Ephemeris data and expertise were provided by W.M.F.\ and R.S.P.