### 2020-01-04 onepager/main.tex

\documentclass[USletter,11pt]{article}
\usepackage[top=1.0in, bottom=1.0in, left=1.0in, right=1.0in,headheight=0.4in]{geometry}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{pslatex} %Times New Roman
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{etaremune}
\usepackage{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{wrapfig}
\usepackage[skip=2pt]{caption}
\usepackage{ifthen}
\usepackage{fancyhdr}

\AtBeginDocument{\RequirePackage{lmodern, times}} %converts math back to Computer Modern

\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\DeclareGraphicsRule{.ai}{pdf}{.ai}{}

\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

% Header graphics
\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\fancyheadoffset{0.5cm}
\fancyhead[L]{\ifthenelse{\value{page}=1}{\includegraphics[width=0.12\textwidth]{nanograv_pfc_logo.png}}{}}
\fancyhead[R]{\ifthenelse{\value{page}=1}{\includegraphics[width=0.1\textwidth]{NSF_4-Color_vector_Logo.jpg}}{}}

\setlength{\parskip}{0.5ex} %2ex
\setlength{\columnsep}{5cm}