### 2020-01-22 paper-apj/authors.tex
Submitted ApJ version

% named authors: M.~Vallisneri, S.~R.~Taylor, J.~Simon, W.~M.~Folkner, R.~S.~Park, C.~Cutler, J.~A.~Ellis, T.~J.~W.~Lazio, S.~J.~Vigeland

\author{
%The NANOGrav Collaboration:
M.~Vallisneri\altaffilmark{1,2},
S.~R.~Taylor\altaffilmark{3,1,2},
J.~Simon\altaffilmark{1,2},
W.~M.~Folkner\altaffilmark{1},
R.~S.~Park\altaffilmark{1},
C.~Cutler\altaffilmark{1,2},
J.~A.~Ellis\altaffilmark{4},
T.~J.~W.~Lazio\altaffilmark{1,2},
S.~J.~Vigeland\altaffilmark{5},
K.~Aggarwal\altaffilmark{6,7},
Z.~Arzoumanian\altaffilmark{8},
P.~T.~Baker\altaffilmark{9,7},
A.~Brazier\altaffilmark{10,11},
P.~R.~Brook\altaffilmark{6,7},
S.~Burke-Spolaor\altaffilmark{6,7},
S.~Chatterjee\altaffilmark{10},
J.~M.~Cordes\altaffilmark{10},
N.~J.~Cornish\altaffilmark{12},
F.~Crawford\altaffilmark{13},
H.~T.~Cromartie\altaffilmark{14},
K.~Crowter\altaffilmark{15},
M.~DeCesar\altaffilmark{$\dagger$,16},
P.~B.~Demorest\altaffilmark{17},
T.~Dolch\altaffilmark{18},
R.~D.~Ferdman\altaffilmark{19},
E.~C.~Ferrara\altaffilmark{20},
E.~Fonseca\altaffilmark{21},
N.~Garver-Daniels\altaffilmark{6,7},
P.~Gentile\altaffilmark{6,7},
D.~Good\altaffilmark{15},
J.~S.~Hazboun\altaffilmark{$\dagger$,22},
A.~M.~Holgado\altaffilmark{23},
E.~A.~Huerta\altaffilmark{23},
K.~Islo\altaffilmark{5},
R.~Jennings\altaffilmark{10},
G.~Jones\altaffilmark{24},
M.~L.~Jones\altaffilmark{5},
D.~L.~Kaplan\altaffilmark{5},
L.~Z.~Kelley\altaffilmark{25},
J.~S.~Key\altaffilmark{22},
M.~T.~Lam\altaffilmark{26,27,6,7},
L.~Levin\altaffilmark{28},
D.~R.~Lorimer\altaffilmark{6,7},
J.~Luo\altaffilmark{21},
R.~S.~Lynch\altaffilmark{29},
D.~R.~Madison\altaffilmark{$\dagger$,6,7},
M.~A.~McLaughlin\altaffilmark{6,7},
S.~T.~McWilliams\altaffilmark{6,7},
C.~M.~F.~Mingarelli\altaffilmark{30,31},
C.~Ng\altaffilmark{32},
D.~J.~Nice\altaffilmark{16},
T.~T.~Pennucci\altaffilmark{$\dagger$,33},
N.~S.~Pol\altaffilmark{6,7},
S.~M.~Ransom\altaffilmark{14,34},
P.~S.~Ray\altaffilmark{35},
X.~Siemens\altaffilmark{36,5},
R.~Spiewak\altaffilmark{5,37},
I.~H.~Stairs\altaffilmark{15},
D.~R.~Stinebring\altaffilmark{38},
K.~Stovall\altaffilmark{$\dagger$,17},
J.~K.~Swiggum\altaffilmark{$\dagger$,5},
R.~van~Haasteren\altaffilmark{1},
C.~A.~Witt\altaffilmark{6,7},
W.~W.~Zhu\altaffilmark{39}\\
(The NANOGrav Collaboration)\altaffilmark{$\star$}}

\affil{$\star$Author order alphabetical by surname}

\affil{$^{1}$Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\affil{$^{2}$Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\affil{$^{3}$Department of Physics and Astronomy, Vanderbilt University, 2301 Vanderbilt Place, Nashville, TN 37235, USA}
\affil{$^{4}$Infinia ML, 202 Rigsbee Avenue, Durham, NC 27701, USA}
\affil{$^{5}$Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\affil{$^{6}$Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affil{$^{7}$Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\affil{$^{8}$X-Ray Astrophysics Laboratory, NASA Goddard Space Flight Center, Code 662, Greenbelt, MD 20771, USA}
\affil{$^{9}$Department of Physics and Astronomy, Widener University, One University Place, Chester, PA 19013, USA}
\affil{$^{10}$Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}
\affil{$^{11}$Cornell Center for Advanced Computing, Ithaca, NY 14853, USA}
\affil{$^{12}$Department of Physics, Montana State University, Bozeman, MT 59717, USA}
\affil{$^{13}$Department of Physics and Astronomy, Franklin \& Marshall College, P.O.~Box 3003, Lancaster, PA 17604, USA}
\affil{$^{14}$University of Virginia, Department of Astronomy, P.O.~Box 400325, Charlottesville, VA 22904, USA}
\affil{$^{15}$Department of Physics and Astronomy, University of British Columbia, 6224 Agricultural Road, Vancouver, BC V6T 1Z1, Canada}
\affil{$^{16}$Department of Physics, Lafayette College, Easton, PA 18042, USA}
\affil{$^{17}$National Radio Astronomy Observatory, 1003 Lopezville Rd., Socorro, NM 87801, USA}
\affil{$^{18}$Department of Physics, Hillsdale College, 33 E.~College Street, Hillsdale, MI 49242, USA}
\affil{$^{19}$Department of Physics, University of East Anglia, Norwich, UK}
\affil{$^{20}$NASA Goddard Space Flight Center, Greenbelt, MD 20771, USA}
\affil{$^{21}$Department of Physics, McGill University, 3600  University St., Montreal, QC H3A 2T8, Canada}
\affil{$^{22}$University of Washington Bothell, 18115 Campus Way NE, Bothell, WA 98011, USA}
\affil{$^{23}$NCSA and Department of Astronomy, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}
\affil{$^{24}$Department of Physics, Columbia University, New York, NY 10027, USA}
\affil{$^{25}$Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), Northwestern University, Evanston, IL 60208}
\affil{$^{26}$School of Physics and Astronomy, Rochester Institute of Technology, Rochester, NY 14623, USA}
\affil{$^{27}$Laboratory for Multiwavelength Astronomy, Rochester Institute of Technology, Rochester, NY 14623, USA}
\affil{$^{28}$Jodrell Bank Centre for Astrophysics, University of Manchester, Manchester, M13 9PL, United Kingdom}
\affil{$^{29}$Green Bank Observatory, P.O.~Box 2, Green Bank, WV 24944, USA}
\affil{$^{30}$Department of Physics, University of Connecticut, 196 Auditorium Road, U-3046, Storrs, CT 06269-3046, USA}
\affil{$^{31}$Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York, NY 10010, USA}
\affil{$^{32}$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St. George St., Toronto, ON M5S 3H4, Canada}
\affil{$^{33}$Hungarian Academy of Sciences MTA-ELTE Extragalactic Astrophysics Research Group, Institute of Physics, E\"{o}tv\"{o}s Lor\'{a}nd University, P\'{a}zm\'{a}ny {P.s.} 1/A, 1117 Budapest, Hungary}
\affil{$^{34}$National Radio Astronomy Observatory, 520 Edgemont Road, Charlottesville, VA 22903, USA}
\affil{$^{35}$Naval Research Laboratory, Washington DC 20375, USA}
\affil{$^{36}$Department of Physics, Oregon State University, Corvallis, OR 97331, USA}
\affil{$^{37}$Centre for Astrophysics and Supercomputing, Swinburne University of Technology, PO Box 218, Hawthorn, VIC 3122, Australia}
\affil{$^{38}$Department of Physics and Astronomy, Oberlin College, Oberlin, OH 44074, USA}
\affil{$^{39}$CAS Key Laboratory of FAST, Chinese Academy of Science, Beijing 100101, China}

\affil{$^{\dagger}$NANOGrav Physics Frontiers Center Postdoctoral Fellow}
% \affil{$^{\ddagger}$ currently at Microsoft Corporation}
\email[Corresponding author email: ]{Michele.Vallisneri@jpl.nasa.gov}